Analyzing the influence of Selection on Genetic Programming's Generalization ability in Symbolic Regression

### 2022-06-27 docs/rmd/presentation/slides.tex
autogit

\begin{frame}{Research Question}
\protect\hypertarget{research-question-1}{}
\begin{itemize}
\tightlist
\item
  Does the usage of \(\epsilon\)-lexicase parent selection influence the
  generalization behavior of genetic programming in symbolic regression
  if compared to tournament selection?
\end{itemize}

\begin{block}{Statistical Testing Strategy}
\protect\hypertarget{statistical-testing-strategy}{}
\begin{enumerate}
\tightlist
\item
  Differences in average fitness for both algorithms on both datasets?
\item
  Differences in fitness for training and testing data?
\end{enumerate}
\end{block}
\end{frame}

@@

\caption{\label{tab:unnamed-chunk-2}Mean Error - P-Values (MWU)}
\centering
\fontsize{4}{6}\selectfont
\begin{tabular}[t]{l|r|r|r|r}
\hline
X & tournament\_training\_errors & tournament\_testing\_errors & elexicase\_training\_errors & elexicase\_testing\_errors\\
\hline
tournament\_training\_errors & 1.000 & 0.309 & 0.000 & 0.000\\
\hline
tournament\_testing\_errors & 0.309 & 1.000 & 0.002 & 0.000\\
\hline
elexicase\_training\_errors & 0.000 & 0.002 & 1.000 & 0.257\\
\hline
elexicase\_testing\_errors & 0.000 & 0.000 & 0.257 & 1.000\\
\hline
\end{tabular}
\end{table}
\end{frame}

@@

\caption{\label{tab:unnamed-chunk-3}Evolutionary Parameters}
\centering
\begin{tabular}[t]{l|l}
\hline
Parameter & Value\\
\hline
Population Size & 500\\
\hline
Number of Generations & 100\\
\hline
Mutation Rate & 20\%\\
\hline
Crossover Rate & 80\%\\
\hline
Tournament Size & 3\\
\hline
Epsilon selection & automatic\\
\hline
Elite Size & 0\\
\hline
\end{tabular}
\end{table}
\end{frame}

@@

\caption{\label{tab:unnamed-chunk-4}Function Set}
\centering
\begin{tabular}[t]{l|r}
\hline
Function & Arity\\
\hline
Addition & 2\\
\hline
Subtraction & 2\\
\hline
Multiplication & 2\\
\hline
Negation & 1\\
\hline
Sine & 1\\
\hline
Cosine & 1\\
\hline
Protected Division & 2\\
\hline
Protected Natural Logarithm & 1\\
\hline
Protected Square Root & 1\\
\hline
\end{tabular}
\end{table}
\framebreak

@@

\caption{\label{tab:unnamed-chunk-5}Terminal Set}
\centering
\begin{tabular}[t]{l|l}
\hline
Terminal & Description\\
\hline
X1 & Relative Compactness\\
\hline
X2 & Surface Area\\
\hline
X3 & Wall Area\\
\hline
X4 & Roof Area\\
\hline
X5 & Overall Height\\
\hline
X6 & Orientation\\
\hline
X7 & Glazing Area\\
\hline
X8 & Glazing Area Distribution\\
\hline
random\_int & Ephemeral Constant (integer)\\
\hline
random\_float & Ephemeral Constant(float)\\
\hline
\end{tabular}
\end{table}
\end{frame}

>>>>

\begin{frame}{Hypothesis Testing}
\protect\hypertarget{hypothesis-testing}{}
\textbf{Does the usage of \(\epsilon\)-lexicase parent selection
influence the generalization behavior of genetic programming in symbolic
regression if compared to tournament selection?}

\begin{enumerate}
\tightlist
\item
  Test for differences in average fitness between both algorithms
\item
  Test for differences in average fitness between training and testing
  data
\end{enumerate}
\end{frame}

@@

\caption{\label{tab:pvals_fitness}Mean Error - P-Values (MWU)}
\centering
\fontsize{4}{6}\selectfont
\begin{tabular}[t]{l|r|r|r|r}
\hline
X & tournament\_training\_errors & tournament\_testing\_errors & elexicase\_training\_errors & elexicase\_testing\_errors\\
\hline
tournament\_training\_errors & 1.000 & 0.309 & 0.000 & 0.000\\
\hline
tournament\_testing\_errors & 0.309 & 1.000 & 0.002 & 0.000\\
\hline
elexicase\_training\_errors & 0.000 & 0.002 & 1.000 & 0.257\\
\hline
elexicase\_testing\_errors & 0.000 & 0.000 & 0.257 & 1.000\\
\hline
\end{tabular}
\end{table}
\end{frame}

@@

\caption{\label{tab:ea_config}Evolutionary Parameters}
\centering
\begin{tabular}[t]{l|l}
\hline
Parameter & Value\\
\hline
Population Size & 500\\
\hline
Number of Generations & 100\\
\hline
Mutation Rate & 20\%\\
\hline
Crossover Rate & 80\%\\
\hline
Tournament Size & 3\\
\hline
Epsilon selection & automatic\\
\hline
Elite Size & 0\\
\hline
\end{tabular}
\end{table}
\end{frame}

@@

\caption{\label{tab:primitive_table}Function Set}
\centering
\begin{tabular}[t]{l|r}
\hline
Function & Arity\\
\hline
Addition & 2\\
\hline
Subtraction & 2\\
\hline
Multiplication & 2\\
\hline
Negation & 1\\
\hline
Sine & 1\\
\hline
Cosine & 1\\
\hline
Protected Division & 2\\
\hline
Protected Natural Logarithm & 1\\
\hline
Protected Square Root & 1\\
\hline
\end{tabular}
\end{table}
\framebreak

@@

\caption{\label{tab:function_table}Terminal Set}
\centering
\begin{tabular}[t]{l|l}
\hline
Terminal & Description\\
\hline
X1 & Relative Compactness\\
\hline
X2 & Surface Area\\
\hline
X3 & Wall Area\\
\hline
X4 & Roof Area\\
\hline
X5 & Overall Height\\
\hline
X6 & Orientation\\
\hline
X7 & Glazing Area\\
\hline
X8 & Glazing Area Distribution\\
\hline
random\_int & Ephemeral Constant (integer)\\
\hline
random\_float & Ephemeral Constant(float)\\
\hline
\end{tabular}
\end{table}
\end{frame}
